---
title: "Models"
author: "Antonio J. Pérez-Luque"
date: "2022-10-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
```{r, message=FALSE}
source("code/load_pkgs.R")
source("code/aux_functions.R")
```

```{r setup, include=FALSE}
## Global options
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = TRUE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               autodep = TRUE)
```

```{r data-raw} 
d_raw <- read_csv(here::here("data/cleaned_data.csv"), show_col_types = FALSE)

selected_vars <- c("sr.spr",
                   "sr.sum",
                   "sr.win",
                   "cic",
                   "fosforo",
                   "ph", 
                   "potasio",
                   "nitrogeno")
```

The potential selected variables are (see [`compare_commons`](compare_commons.Rmd)): 

```{r variables}
d_raw |> dplyr::select(one_of(selected_vars)) |> names()
```

```{r data}
d <- d_raw |> 
  dplyr::select(one_of(selected_vars),
                commons, 
                transh,
                parcela,
                quadrat,
                shannon,
                fitovol_dm3,
                rec_total_porc,
                rec_vegetal_porc,
                riqueza,
                tasa_cons) %>% 
  mutate(parcela = as.factor(parcela))
```

```{r prepare-covariates}
covariates <- c(selected_vars, 
                "commons",
                "transh")    

full_covariates <- glue::glue_collapse(covariates, sep = "+")

random_part <- "(1 | parcela)"
```

```{r data-plot}
data_mean <- read_csv(here::here("data/data_mean.csv"), show_col_types = FALSE)
```


## Richness

```{r modeliza-richness}
yvar <- "riqueza"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-richness}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- lm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux) 
```

```{r checkmodel-richness, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for richness. 

### Model selection 

- Perform a multimodel selection

```{r multimodel-richness, message=FALSE}
yvar <- "riqueza"
covariates <- glue::glue_collapse(selected_vars, sep = " + ")
always <- "+ commons + transh" 
f <- "riqueza ~ sr.spr + sr.sum + cic + fosforo + ph + potasio + nitrogeno"

lmer.glmulti <- function(formula, data, random = "", always = "", ...) {
    newf <- formula
    newf[[3]] <- substitute(f + a + r,
                            list(f = newf[[3]], 
                                 r = reformulate(random)[[2]],
                                 a = reformulate(always)[[2]]))
    lmer(newf, data = data,
         REML = FALSE, ...)
}

#https://yuzar-blog.netlify.app/posts/2022-05-31-glmulti/

set.seed(1234)
mm <- glmulti(
  f, 
  always = c("commons","transh"), 
  random =  random_part,
  data = d,
  method = "ga",
  deltaM = 0.5,
  maxsize = 5, 
  fitfunction = lmer.glmulti,
  marginality = FALSE, 
  plotty = FALSE,
  report = FALSE,
  level = 1) # 1 without interaction


# Best models (delta AIC <= 2)
top <- weightable(mm)
top <- top[top$aic <= (min(top$aic) + 2),]
```

- Get the best models (maximum `r nrow(top)` models ($\Delta_{AIC}<2$))

```{r bestmodels-richness}
# Get the ten best models 
maxmodels <- nrow(top)

for (i in 1:maxmodels) { 
  mod <- mm@objects[[i]]
  name_model <- paste0("best_mod_", i)
  assign(name_model, mod)}


# Get the table 
modelos <- paste0("best_mod_", 1:maxmodels)
lista_model <- mget(modelos, envir = globalenv())

lista_modelos <- model.sel(lista_model, rank = AIC) %>% 
  rownames_to_column("model")
  
var_formulas <- data.frame(
  model = paste0("best_mod_", 1:maxmodels),
  formula = 'NA'
)

for (i in 1:maxmodels) {
  # Ojo no incluyo la parte aleatoria que esta en todos. Por eso indico el [[2]] 
  var_formulas[i, "formula"] <- deparse(eval(parse(text = paste0("formula(best_mod_",i, ")[[3]]")))[[2]])[[1]]
}

lista_modelos <- lista_modelos %>% 
  inner_join(var_formulas) %>% 
  as.data.frame() %>% 
  relocate(formula, .after = model) %>% 
  dplyr::select(-model)

lista_modelos %>% 
  write.csv(here::here(paste0("output/best_models_", yvar, ".csv")), 
            na = "") 

lista_modelos_aic <- lista_modelos %>% 
  dplyr::select(formula, df, logLik, AIC, delta, weight) %>% 
  mutate_at(3:4, ~round(.,2)) %>% 
  mutate_at(5:6, ~round(.,3)) 

lista_modelos_aic %>% 
  write.csv(here::here(paste0("output/best_models_aic_", yvar, ".csv")), 
            na = "")
```

```{r bestmodelsAIC-richness}
lista_modelos_aic %>% 
  kbl(caption = paste0("Model selection of ", yvar)) %>% 
  kable_styling()
```

- Explore the variable importance.


```{r varImp-richness}
# Variable Importance 
plot(mm, type = "s", sub = yvar, font.sub = 2, cex.sub = 1.5)
abline(v = 0.5, col = "blue")
```

- Select the best and most parsimonious model 

```{r model-richenss}
bm <- lmer(
  riqueza ~ sr.sum + cic + ph + commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

assign(paste0('table_performance', yvar), (performance(bm) |> mutate(yvar = yvar))) 

```

```{r anova-richness}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-richness}
tabla_coef(bm)
```

```{r r2-richness}
assign(paste0('table_glance', yvar), (glance(bm) |> mutate(yvar = yvar))) 
```


```{r clean-richness}
rm(list = apropos("best_mod"))
rm(lista_model)
```

### Posthoc 
```{r posthoc-richness}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-richness}
name_ylab <- "Richness (species number)"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Diversity (shannon)

```{r modeliza-shannon}
yvar <- "shannon"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-shannon}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- lm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux) 
```

```{r checkmodel-shannon, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for diversity (Shannon index) 

### Model selection 

- Perform a multimodel selection

```{r multimodel-shannon, message=FALSE}
yvar <- "shannon"
covariates <- glue::glue_collapse(selected_vars, sep = " + ")
always <- "+ commons + transh" 
f <- "shannon ~ sr.spr + sr.sum + cic + fosforo + ph + potasio + nitrogeno"

lmer.glmulti <- function(formula, data, random = "", always = "", ...) {
    newf <- formula
    newf[[3]] <- substitute(f + a + r,
                            list(f = newf[[3]], 
                                 r = reformulate(random)[[2]],
                                 a = reformulate(always)[[2]]))
    lmer(newf, data = data,
         REML = FALSE, ...)
}

#https://yuzar-blog.netlify.app/posts/2022-05-31-glmulti/

set.seed(1234)
mm <- glmulti(
  f, 
  always = c("commons","transh"), 
  random =  random_part,
  data = d,
  method = "ga",
  deltaM = 0.5,
  maxsize = 5, 
  fitfunction = lmer.glmulti,
  marginality = FALSE, 
  plotty = FALSE,
  report = FALSE,
  level = 1) # 1 without interaction


# Best models (delta AIC <= 2)
top <- weightable(mm)
top <- top[top$aic <= (min(top$aic) + 2),]
```

- Get the best models (maximum `r nrow(top)` models ($\Delta_{AIC}<2$))

```{r bestmodels-shannon}
# Get the ten best models 
maxmodels <- nrow(top)

for (i in 1:maxmodels) { 
  mod <- mm@objects[[i]]
  name_model <- paste0("best_mod_", i)
  assign(name_model, mod)}


# Get the table 
modelos <- paste0("best_mod_", 1:maxmodels)
lista_model <- mget(modelos, envir = globalenv())

lista_modelos <- model.sel(lista_model, rank = AIC) %>% 
  rownames_to_column("model")
  
var_formulas <- data.frame(
  model = paste0("best_mod_", 1:maxmodels),
  formula = 'NA'
)

for (i in 1:maxmodels) {
  # Ojo no incluyo la parte aleatoria que esta en todos. Por eso indico el [[2]] 
  var_formulas[i, "formula"] <- deparse(eval(parse(text = paste0("formula(best_mod_",i, ")[[3]]")))[[2]])[[1]]
}

lista_modelos <- lista_modelos %>% 
  inner_join(var_formulas) %>% 
  as.data.frame() %>% 
  relocate(formula, .after = model) %>% 
  dplyr::select(-model)

lista_modelos %>% 
  write.csv(here::here(paste0("output/best_models_", yvar, ".csv")), 
            na = "") 

lista_modelos_aic <- lista_modelos %>% 
  dplyr::select(formula, df, logLik, AIC, delta, weight) %>% 
  mutate_at(3:4, ~round(.,2)) %>% 
  mutate_at(5:6, ~round(.,3)) 

lista_modelos_aic %>% 
  write.csv(here::here(paste0("output/best_models_aic_", yvar, ".csv")), 
            na = "")
```

```{r bestmodelsAIC-shannon}
lista_modelos_aic %>% 
  kbl(caption = paste0("Model selection of ", yvar)) %>% 
  kable_styling() 
```

- Explore the variable importance.[^1]

[^1]: Sum of the weights/probabilities for the models in which the variable appears. So, a variable that shows up in lots of models with large weights will receive a high importance value. Some authors refers as 0.8 cutoff and others 0.5 cutoff

```{r varImp-shannon}
# Variable Importance 
plot(mm, type = "s", sub = yvar, font.sub = 2, cex.sub = 1.5)
abline(v = 0.5, col = "blue")
```

- Select the best and most parsimonious model 

```{r model-shannon}
bm <- lmer(
  shannon ~ cic + nitrogeno + commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

assign(paste0('table_performance', yvar), (performance(bm) |> mutate(yvar = yvar))) 


```

```{r anova-shannon}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-shannon}
tabla_coef(bm)
```

```{r r2-shannon}
assign(paste0('table_glance', yvar), (glance(bm) |> mutate(yvar = yvar))) 
```


```{r clean-shannon}
rm(list = apropos("best_mod"))
rm(lista_model)
```

### Posthoc 
```{r posthoc-shannon}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-shannon}
name_ylab <- "Shannon index"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```


## Total coverage (rec_total_porc) 

```{r modeliza-rec_total}
yvar <- "rec_total_porc"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-rec_total}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-rec_total, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for Total coverage 

### Model selection 

- Perform a multimodel selection

```{r multimodel-rec_total, message=FALSE}
yvar <- "rec_total_porc"
covariates <- glue::glue_collapse(selected_vars, sep = " + ")
always <- "+ commons + transh" 
f <- "rec_total_porc ~ sr.spr + sr.sum + cic + fosforo + ph + potasio + nitrogeno"

lmer.glmulti <- function(formula, data, random = "", always = "", ...) {
    newf <- formula
    newf[[3]] <- substitute(f + a + r,
                            list(f = newf[[3]], 
                                 r = reformulate(random)[[2]],
                                 a = reformulate(always)[[2]]))
    lmer(newf, data = data,
         REML = FALSE, ...)
}

#https://yuzar-blog.netlify.app/posts/2022-05-31-glmulti/

set.seed(1234)
mm <- glmulti(
  f, 
  always = c("commons","transh"), 
  random =  random_part,
  data = d,
  method = "ga",
  deltaM = 0.5,
  maxsize = 5, 
  fitfunction = lmer.glmulti,
  marginality = FALSE, 
  plotty = FALSE,
  report = FALSE,
  level = 1) # 1 without interaction


# Best models (delta AIC <= 2)
top <- weightable(mm)
top <- top[top$aic <= (min(top$aic) + 2),]
```

- Get the best models (maximum `r nrow(top)` models ($\Delta_{AIC}<2$))

```{r bestmodels-rec_total}
# Get the ten best models 
maxmodels <- nrow(top)

for (i in 1:maxmodels) { 
  mod <- mm@objects[[i]]
  name_model <- paste0("best_mod_", i)
  assign(name_model, mod)}


# Get the table 
modelos <- paste0("best_mod_", 1:maxmodels)
lista_model <- mget(modelos, envir = globalenv())

lista_modelos <- model.sel(lista_model, rank = AIC) %>% 
  rownames_to_column("model")
  
var_formulas <- data.frame(
  model = paste0("best_mod_", 1:maxmodels),
  formula = 'NA'
)

for (i in 1:maxmodels) {
  # Ojo no incluyo la parte aleatoria que esta en todos. Por eso indico el [[2]] 
  var_formulas[i, "formula"] <- deparse(eval(parse(text = paste0("formula(best_mod_",i, ")[[3]]")))[[2]])[[1]]
}

lista_modelos <- lista_modelos %>% 
  inner_join(var_formulas) %>% 
  as.data.frame() %>% 
  relocate(formula, .after = model) %>% 
  dplyr::select(-model)

lista_modelos %>% 
  write.csv(here::here(paste0("output/best_models_", yvar, ".csv")), 
            na = "") 

lista_modelos_aic <- lista_modelos %>% 
  dplyr::select(formula, df, logLik, AIC, delta, weight) %>% 
  mutate_at(3:4, ~round(.,2)) %>% 
  mutate_at(5:6, ~round(.,3)) 

lista_modelos_aic %>% 
  write.csv(here::here(paste0("output/best_models_aic_", yvar, ".csv")), 
            na = "")
```

```{r bestmodelsAIC-rec_total}
lista_modelos_aic %>% 
  kbl(caption = paste0("Model selection of ", yvar)) %>% 
  kable_styling() 
```

- Explore the variable importance.


```{r varImp-rec_total}
# Variable Importance 
plot(mm, type = "s", sub = yvar, font.sub = 2, cex.sub = 1.5)
abline(v = 0.5, col = "blue")
```

- Select the best and most parsimonious model 

```{r model-rec_total}
bm <- lmer(
  rec_total_porc ~ sr.spr + sr.sum + ph + potasio + commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

assign(paste0('table_performance', yvar), (performance(bm) |> mutate(yvar = yvar))) 


```

```{r anova-rec_total}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-rec_total}
tabla_coef(bm)
```

```{r r2-rec_total}
assign(paste0('table_glance', yvar), (glance(bm) |> mutate(yvar = yvar))) 
```

```{r clean-rec_total}
rm(list = apropos("best_mod"))
rm(lista_model)
```

### Posthoc 
```{r posthoc-rec_total}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-rec_total}
name_ylab <- "Total Coverage (%)"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Phytovolumen (fitovol_dm3)

```{r modeliza-fitvol}
yvar <- "fitovol_dm3"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-fitvol}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-fitvol, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for Phytovolumen 

### Model selection 

- Perform a multimodel selection

```{r multimodel-fitvol, message=FALSE}
yvar <- "fitovol_dm3"
covariates <- glue::glue_collapse(selected_vars, sep = " + ")
always <- "+ commons + transh" 
f <- "fitovol_dm3 ~ sr.spr + sr.sum + cic + fosforo + ph + potasio + nitrogeno"

lmer.glmulti <- function(formula, data, random = "", always = "", ...) {
    newf <- formula
    newf[[3]] <- substitute(f + a + r,
                            list(f = newf[[3]], 
                                 r = reformulate(random)[[2]],
                                 a = reformulate(always)[[2]]))
    lmer(newf, data = data,
         REML = FALSE, ...)
}

#https://yuzar-blog.netlify.app/posts/2022-05-31-glmulti/

set.seed(1234)
mm <- glmulti(
  f, 
  always = c("commons","transh"), 
  random =  random_part,
  data = d,
  method = "ga",
  deltaM = 0.5,
  maxsize = 5, 
  fitfunction = lmer.glmulti,
  marginality = FALSE, 
  plotty = FALSE,
  report = FALSE,
  level = 1) # 1 without interaction


# Best models (delta AIC <= 2)
top <- weightable(mm)
top <- top[top$aic <= (min(top$aic) + 2),]
```

- Get the best models (maximum `r nrow(top)` models ($\Delta_{AIC}<2$))

```{r bestmodels-fitovol}
# Get the ten best models 
maxmodels <- nrow(top)

for (i in 1:maxmodels) { 
  mod <- mm@objects[[i]]
  name_model <- paste0("best_mod_", i)
  assign(name_model, mod)}


# Get the table 
modelos <- paste0("best_mod_", 1:maxmodels)
lista_model <- mget(modelos, envir = globalenv())

lista_modelos <- model.sel(lista_model, rank = AIC) %>% 
  rownames_to_column("model")
  
var_formulas <- data.frame(
  model = paste0("best_mod_", 1:maxmodels),
  formula = 'NA'
)

for (i in 1:maxmodels) {
  # Ojo no incluyo la parte aleatoria que esta en todos. Por eso indico el [[2]] 
  var_formulas[i, "formula"] <- deparse(eval(parse(text = paste0("formula(best_mod_",i, ")[[3]]")))[[2]])[[1]]
}

lista_modelos <- lista_modelos %>% 
  inner_join(var_formulas) %>% 
  as.data.frame() %>% 
  relocate(formula, .after = model) %>% 
  dplyr::select(-model)

lista_modelos %>% 
  write.csv(here::here(paste0("output/best_models_", yvar, ".csv")), 
            na = "") 

lista_modelos_aic <- lista_modelos %>% 
  dplyr::select(formula, df, logLik, AIC, delta, weight) %>% 
  mutate_at(3:4, ~round(.,2)) %>% 
  mutate_at(5:6, ~round(.,3)) 

lista_modelos_aic %>% 
  write.csv(here::here(paste0("output/best_models_aic_", yvar, ".csv")), 
            na = "")
```

```{r bestmodelsAIC-fitovol}
lista_modelos_aic %>% 
  kbl(caption = paste0("Model selection of ", yvar)) %>% 
  kable_styling() 
```

- Explore the variable importance.


```{r varImp-fitovol}
# Variable Importance 
plot(mm, type = "s", sub = yvar, font.sub = 2, cex.sub = 1.5)
abline(v = 0.5, col = "blue")
```

- Select the best and most parsimonious model 

```{r model-fitovol}
bm <- lmer(
  fitovol_dm3 ~ sr.sum + cic + commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

assign(paste0('table_performance', yvar), (performance(bm) |> mutate(yvar = yvar))) 


```

```{r anova-fitvol}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-fitvol}
tabla_coef(bm)
```

```{r r2-fitvol}
assign(paste0('table_glance', yvar), (glance(bm) |> mutate(yvar = yvar))) 
```


```{r clean-fitvol}
rm(list = apropos("best_mod"))
rm(lista_model)
```

### Posthoc 
```{r posthoc-fitovol}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-fitovol}
name_ylab <- expression(Fitovolumen~(m^3))
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Consumption rate(tasa_cons)

```{r modeliza-tasacons}
yvar <- "tasa_cons"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-tasacons}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-tasacons, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

 
### Model selection 

- Perform a multimodel selection

```{r multimodel-tasacons, message=FALSE}
yvar <- "tasa_cons"
covariates <- glue::glue_collapse(selected_vars, sep = " + ")
always <- "+ commons + transh" 
f <- "tasa_cons ~ sr.spr + sr.sum + cic + fosforo + ph + potasio + nitrogeno"

lmer.glmultiGamma <- function(formula, data, random = "", always = "", ...) {
    newf <- formula
    newf[[3]] <- substitute(f + a + r,
                            list(f = newf[[3]], 
                                 r = reformulate(random)[[2]],
                                 a = reformulate(always)[[2]]))
    glmer(newf, data = data, ...)
}

#https://yuzar-blog.netlify.app/posts/2022-05-31-glmulti/

set.seed(1234)
mm <- glmulti(
  f, 
  always = c("commons","transh"), 
  random =  random_part,
  family = Gamma(link = "log"),
  data = d,
  method = "ga",
  deltaM = 0.5,
  maxsize = 5, 
  fitfunction = lmer.glmultiGamma,
  marginality = FALSE, 
  plotty = FALSE,
  report = FALSE,
  level = 1) # 1 without interaction


# Best models (delta AIC <= 2)
top <- weightable(mm)
top <- top[top$aic <= (min(top$aic) + 2),]
```

- Get the best models (maximum `r nrow(top)` models ($\Delta_{AIC}<2$))

```{r bestmodels-tasacons}
# Get the ten best models 
maxmodels <- nrow(top)

for (i in 1:maxmodels) { 
  mod <- mm@objects[[i]]
  name_model <- paste0("best_mod_", i)
  assign(name_model, mod)}


# Get the table 
modelos <- paste0("best_mod_", 1:maxmodels)
lista_model <- mget(modelos, envir = globalenv())

lista_modelos <- model.sel(lista_model, rank = AIC) %>% 
  rownames_to_column("model")
  
var_formulas <- data.frame(
  model = paste0("best_mod_", 1:maxmodels),
  formula = 'NA'
)

for (i in 1:maxmodels) {
  # Ojo no incluyo la parte aleatoria que esta en todos. Por eso indico el [[2]] 
  var_formulas[i, "formula"] <- deparse(eval(parse(text = paste0("formula(best_mod_",i, ")[[3]]")))[[2]])[[1]]
}

lista_modelos <- lista_modelos %>% 
  inner_join(var_formulas) %>% 
  as.data.frame() %>% 
  relocate(formula, .after = model) %>% 
  dplyr::select(-model)

lista_modelos %>% 
  write.csv(here::here(paste0("output/best_models_", yvar, ".csv")), 
            na = "") 

lista_modelos_aic <- lista_modelos %>% 
  dplyr::select(formula, df, logLik, AIC, delta, weight) %>% 
  mutate_at(3:4, ~round(.,2)) %>% 
  mutate_at(5:6, ~round(.,3)) 

lista_modelos_aic %>% 
  write.csv(here::here(paste0("output/best_models_aic_", yvar, ".csv")), 
            na = "")
```

```{r bestmodelsAIC-tasacons}
lista_modelos_aic %>% 
  kbl(caption = paste0("Model selection of ", yvar)) %>% 
  kable_styling() 
```

- Explore the variable importance.


```{r varImp-tasacons}
# Variable Importance 
plot(mm, type = "s", sub = yvar, font.sub = 2, cex.sub = 1.5)
abline(v = 0.5, col = "blue")
```

- Select the best and most parsimonious model 

```{r model-tasacons}
bm <- lmer(
  tasa_cons ~ 1 + sr.sum + cic + fosforo + commons + transh + commons:transh + (1 | parcela), data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

assign(paste0('table_performance', yvar), (performance(bm) |> mutate(yvar = yvar))) 

```

```{r anova-tasacons}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```


```{r coef-tasacons}
tabla_coef(bm)
```

```{r r2-tasacons}
assign(paste0('table_glance', yvar), (glance(bm) |> mutate(yvar = yvar))) 
```

```{r clean-tasacons}
rm(list = apropos("best_mod"))
rm(lista_model)
```

### Posthoc 
```{r posthoc-tasacons}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-tasacons}
name_ylab <- "Consumption rate"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```


## Tablas 
### ANOVAS

```{r}
anovas <- mget(apropos("anova_"), envir = globalenv()) |> 
  bind_rows() |> 
  remove_rownames() |> 
  relocate(variable, factor) |> 
  mutate_at(3, ~round(.,2)) |> 
  mutate_at(4, ~round(.,3))
```

```{r}
anovas |> 
  kbl(col.names = 
        c("Response", "variables", "F", "p.value")) %>% 
  kable_styling() 
```


### Tablas Glance 

```{r}
tablas_glance <- mget(apropos("table_glance"), envir = globalenv()) |> 
  bind_rows() |> 
  remove_rownames()
```

### Tablas Glance 

```{r}
tablas_R2 <- mget(apropos("table_performance"), envir = globalenv()) |> 
  bind_rows() |> 
  remove_rownames()
```



```{r}
# export tables 
write.csv(anovas, file=here::here("output/anovas_model.csv"), row.names = FALSE)
write.csv(tablas_glance, file=here::here("output/tablas_glance_model.csv"), row.names = FALSE)
write.csv(tablas_R2, file=here::here("output/tablas_R2_model.csv"), row.names = FALSE)
```



















