---
title: "Models Social"
author: "Antonio J. Pérez-Luque"
date: "2022-11-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

## Introduction
```{r, message=FALSE}
source("code/load_pkgs.R")
source("code/aux_functions.R")
```

```{r setup, include=FALSE}
## Global options
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = TRUE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               autodep = TRUE)
```

```{r data-raw} 
d_raw <- read_csv(here::here("data/cleaned_data.csv"), show_col_types = FALSE)

selected_vars <- c("sr.spr",
                   "sr.sum",
                   "sr.win",
                   "cic",
                   "fosforo",
                   "ph", 
                   "potasio",
                   "nitrogeno")
```

The potential selected variables are (see [`compare_commons`](compare_commons.Rmd)): 

```{r variables}
d_raw |> dplyr::select(one_of(selected_vars)) |> names()
```

```{r data}
d <- d_raw |> 
  dplyr::select(one_of(selected_vars),
                commons, 
                transh,
                parcela,
                quadrat,
                shannon,
                fitovol_dm3,
                rec_total_porc,
                rec_vegetal_porc,
                riqueza,
                tasa_cons) %>% 
  mutate(parcela = as.factor(parcela))
```

```{r prepare-covariates}
covariates <- c(selected_vars, 
                "commons",
                "transh")    

full_covariates <- glue::glue_collapse(covariates, sep = "+")

random_part <- "(1 | parcela)"
```

```{r data-plot}
data_mean <- read_csv(here::here("data/data_mean.csv"), show_col_types = FALSE)
```

## Richness

```{r modeliza-richness}
yvar <- "riqueza"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-richness}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- lm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux) 
```

```{r checkmodel-richness, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for richness. 

### Model 

```{r model-richenss}
bm <- lmer(
  riqueza ~ commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()
```

```{r anova-richness}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-richness}
tabla_coef(bm)
```

### Posthoc 
```{r posthoc-richness}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-richness}
name_ylab <- "Richness (species number)"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Diversity (shannon)

```{r modeliza-shannon}
yvar <- "shannon"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-shannon}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- lm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux) 
```

```{r checkmodel-shannon, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for diversity (Shannon index) 

### Model 

```{r model-shannon}
bm <- lmer(
  shannon ~ commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()
```

```{r anova-shannon}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-shannon}
tabla_coef(bm)
```

### Posthoc 
```{r posthoc-shannon}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-shannon}
name_ylab <- "Shannon index"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```


## Total coverage (rec_total_porc) 

```{r modeliza-rec_total}
yvar <- "rec_total_porc"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-rec_total}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-rec_total, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for Total coverage 

### Model  

```{r model-rec_total}
bm <- lmer(
  rec_total_porc ~ commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()
```

```{r anova-rec_total}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-rec_total}
tabla_coef(bm)
```

### Posthoc 
```{r posthoc-rec_total}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-rec_total}
name_ylab <- "Total Coverage (%)"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Phytovolumen (fitovol_dm3)

```{r modeliza-fitvol}
yvar <- "fitovol_dm3"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-fitvol}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-fitvol, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

- It seems that there are no problems of normality or homocedasticity, so we use an LMM for Phytovolumen 

### Model 

```{r model-fitovol}
bm <- lmer(
  fitovol_dm3 ~ commons + transh + commons:transh + (1 | parcela), 
  data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

```

```{r anova-fitvol}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```

```{r coef-fitvol}
tabla_coef(bm)
```

### Posthoc 
```{r posthoc-fitovol}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-fitovol}
name_ylab <- expression(Fitovolumen~(m^3))
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

## Consumption rate(tasa_cons)

```{r modeliza-tasacons}
yvar <- "tasa_cons"
covariates_lmer <- glue::glue_collapse(c(covariates, random_part),
                                       sep = " + ")

f <- reformulate(covariates_lmer, response = yvar)

m <- lmer(glue::glue_collapse(c(f, random_part), " + "),
          data = d)
```

- GLMM or LMM? 

```{r checkGLM-tasacons}
faux <- reformulate(
  glue::glue_collapse(covariates, sep = " + "), response = yvar)

maux <- glm(faux, data = d) 
simulated_resid <- DHARMa::simulateResiduals(maux, n = 1000)
plot(simulated_resid)

plot(performance::check_distribution(maux), panel = FALSE)[[1]] + labs(subtitle = faux)
```

```{r checkmodel-tasacons, fig.height=8}
check_model(maux)
rm(maux, faux, simulated_resid) # remove 
```

 
### Model  


```{r model-tasacons}
bm <- lmer(
  tasa_cons ~ commons + transh + commons:transh + (1 | parcela), data = d
)

performance(bm) %>% 
  kbl() %>% 
  kable_styling()

```

```{r anova-tasacons}
anova(as_lmerModLmerTest(bm))
formatAnova(as.data.frame(anova(as_lmerModLmerTest(bm))), yvar)
```


```{r coef-tasacons}
tabla_coef(bm)
```


### Posthoc 
```{r posthoc-tasacons}
tabla_postHoc_commons(bm)
tabla_postHoc_transh(bm)
tabla_postHoc_int(bm)
tabla_postHoc_transhcommons(bm)
```

```{r plot-tasacons}
name_ylab <- "Consumption rate"
plot_violoin2(df = d, yvar = yvar, name_ylab = name_ylab)
```

```{r}
anovas <- mget(apropos("anova_"), envir = globalenv()) |> 
  bind_rows() |> 
  remove_rownames() |> 
  relocate(variable, factor) |> 
  mutate_at(3, ~round(.,2)) |> 
  mutate_at(4, ~round(.,3))
```

```{r}
anovas |> 
  kbl(col.names = 
        c("Response", "variables", "F", "p.value")) %>% 
  kable_styling() 
```
