---
title: "nmds"
author: "Antonio J. Pérez-Luque"
date: "2022-06-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = TRUE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               autodep = TRUE)
```

```{r pkg}
source(here::here("code/load_pkgs.R"))
source(here::here("code/aux_functions.R"))
```


## Introduction

The aim of the NMDS is to find an optimal ordination of the quadrats based on species abundance-composition to identify any similarity pattern between commons (our sites) and/or type of transhumance. 


### Prepare data  

- An auxiliar field (`id_plot`) was created to join information. We combined the fields `cod_parcela` and `quadrat`. Example: `"CES_1_1"` where `"CES_1"` refers to the `cod_parcela` and `_1` to the quadrat number ($i=1,2,3$). 

- We created two datasets: 

  - a metadata dataset with the metadata information of each of the quadrat 
  - a dataset with the species-composition of each quadrat 
  

```{r read-data}
veg_raw <- readxl::read_excel(path = here::here("data/Datos_para_Antonio.xlsx"), 
                              sheet = "Especies_quadrat") |> 
  janitor::clean_names() |> 
  unite(col = "id", cod_parcela:quadrat, sep = "_", remove = FALSE)

# Dataset A (metadata)
veg_metadatos <- veg_raw |> dplyr::select(localidad:carga_otono)


# Dataset B (species composition)
veg <- veg_raw |> dplyr::select(-c(localidad:tratamiento_humedad,
                                    cod_parcela:carga_otono)) 
```

After the exploration of the composition data, we observed some data identified at different taxonomical level. For instance, some are identified as "Asteracea" and others to species level. So we removed those records identified up to family or up to genus.  


```{r taxa-names}
veg_sel <- veg |> 
  dplyr::select(-ends_with("_sp")) %>% # remove recordds identified up to genus
  dplyr::select(-c(asteraceae,fabaceae)) # remove those two records (up to family identification)

# Create an abbreviate name for the taxa: 3 chracters for genus, 3 for species and 3 for infrataxa. 
# Be careful: Check duplicate names
raw_colnames <- colnames(veg_sel) |>  
  str_remove_all(pattern = "subsp_") |>  
  str_replace_all(pattern = "_",replacement = " ") 

  new_colnames <- c(
    "id",
    fuzzySim::spCodes(raw_colnames[-1],
      nchar.gen = 3, nchar.sp = 3, nchar.ssp = 3
    )
  )

colnames(veg_sel) <- new_colnames
```


## NMDS analysis 

- To avoid convergence issues and high stress (*i.e.* data distortion, [more info](https://aslopubs.onlinelibrary.wiley.com/doi/10.1002/lom3.10257)) values, we used relative abundance (see [this](https://rpubs.com/CPEL/NMDS)). 

```{r} 

# Convertir absolute abundance to relative 
# veg_sel_rel <- cbind(veg_sel[,1], 
#                      decostand(veg_sel[-1], method = "total"))

veg_sel_rel <- decostand(veg_sel[-1], method = "total")
veg_sel_distmat <- vegdist(veg_sel_rel, method = "bray")

veg_sel_distmat <- as.matrix(veg_sel_distmat, labels = TRUE)


nmds <- metaMDS(decostand(veg_sel[-1], method = "total"),
                 distance = "bray", 
                 autotransform = FALSE,
                 maxit = 999, 
                 try = 500, k = 3)

print(nmds)
```


```{r stress-plot}
stressplot(nmds)
```

### Ordination plot 

- **Combined treatments** 
```{r nmds-plot-int, fig.width=9}

# Create df with scores 

nmds_scores <- data.frame(nmds$points, id = veg_sel[,1]) |>  
  inner_join(veg_metadatos) |>
  unite("combinado", localidad:tratamiento_pastor, sep = "-", remove = FALSE) |> 
  rename(commons = localidad, transh = tratamiento_pastor)


theme_nmds <- list(
  geom_point(size = 2.5, alpha = 0.75),
  theme(
    axis.title = element_text(size = 10, face = "bold", colour = "grey30"),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(size = 10, face = "bold", colour = "grey30"),
    legend.text = element_text(size = 9, colour = "grey30")
  )
)

theme_transh <- list(
  stat_ellipse(aes(group=transh), type="norm", level = 0.9, alpha=.5, linetype = 2), 
  scale_color_manual(values = c("#33a02c", "#6a3d9a"))
)

theme_commons <- list(
  stat_ellipse(aes(group=commons), type="norm", level = 0.9, alpha=.5),
  scale_color_manual(values = c("#8c510a","#fdae61", "#01665e"))
)


p12int <- ggplot(nmds_scores, aes(x = MDS1, y = MDS2, color = commons, shape = transh)) +
  theme_nmds 

p13int <- ggplot(nmds_scores, aes(x = MDS1, y = MDS3, color = commons, shape = transh)) +
  theme_nmds

p23int <- ggplot(nmds_scores, aes(x = MDS2, y = MDS3, color = commons, shape = transh)) +
  theme_nmds


p12int + p13int + p23int + plot_layout(guides = "collect") & theme(legend.position = 'top')
```

- **By treatments** 

```{r nmds-plot-both, fig.width=9, fig.height=9}

p12t <- ggplot(nmds_scores, aes(x = MDS1, y = MDS2, shape = transh, colour = transh)) +
  theme_nmds + theme_transh
  
p13t <- ggplot(nmds_scores, aes(x = MDS1, y = MDS3, shape = transh, colour = transh)) +
  theme_nmds + theme_transh

p23t  <- ggplot(nmds_scores, aes(x = MDS2, y = MDS3, shape = transh, colour = transh)) +
  theme_nmds + theme_transh

p12c <- ggplot(nmds_scores, aes(x = MDS1, y = MDS2, shape = commons, colour = commons)) +
  theme_nmds + theme_commons
  
p13c <- ggplot(nmds_scores, aes(x = MDS1, y = MDS3, shape = commons, colour = commons)) +
  theme_nmds + theme_commons

p23c <- ggplot(nmds_scores, aes(x = MDS2, y = MDS3, shape = commons, colour = commons)) +
  theme_nmds + theme_commons

pc <- p12c + p13c + p23c + plot_layout(guides = "collect") & theme(legend.position = 'top')
pt <- p12t + p13t + p23t + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

pc/pt
```


## Jaccard's analysis

```{r}
d <- veg_metadatos |> 
  dplyr::select(
    commons = localidad, 
    transh = tratamiento_pastor, 
    id
  ) |> 
  inner_join(veg_sel)
```

- How is the similarity within plots  (*i.e.* within all quadrats of any treatment combination). So we compare for each treatment the jaccard between quadrats 

- Be careful with the computation of Jaccard in several pkgs. If we used `vegan::vegdist` we are computing the Jaccard distance. So, if we want to compute the similarity Jaccard index we could use `designdist(our_data_frame, method = "a/(a+b+c)", abcd = TRUE)` 



```{r}
df_commons <- d |> dplyr::select(-transh, -id) |> 
  group_by(commons) |> 
  summarise_all(sum)

df_transh <- d |> dplyr::select(-commons, -id) |> 
  group_by(transh) |> 
  summarise_all(sum)

df_ct <- d |> 
  dplyr::select(-transh, -commons) |> 
  mutate(i = substr(id, 1, 3)) |> 
  relocate(i, .after = id) |> 
  dplyr::select(-id) |> 
  group_by(i) |> 
  summarise_all(sum)



ff <- function(mydf, nombres_rows) { 
mydf |> 
  column_to_rownames(var = nombres_rows) |>
  t() |>
  as.data.frame() |>
  rownames_to_column(var = "sp")

}




### Commons 
# C - S 

commons_transpose <- ff(df_commons, nombres_rows = "commons")
tranhs_transpose <- ff(df_transh, nombres_rows = "transh")
ct_transpose <- ff(df_ct, nombres_rows = "i")

myjaccard <- function(df, site1, site2) {

  df <- df[, c("sp", site1, site2)]
  
  names(df) <- c('sp', 'site1', 'site2')
  
s <- df |>
  mutate(J = case_when(
    site1 > 0 & site2 > 0 ~ "a",
    site1 > 0 & site2 == 0 ~ "b",
    site1 == 0 & site2 > 0 ~ "c",
    TRUE ~ "0"
  ))

jcounts <- s %>%
  group_by(J) %>%
  tally()

# Jaccard = J11 / (J10 + J01 + J11)

a <- (jcounts |> filter(J == "a") |> pull(n))
b <- (jcounts |> filter(J == "b") |> pull(n))
c <- (jcounts |> filter(J == "c") |> pull(n))

j <- a / (a + b + c)

return(j)

}


jcp <- myjaccard(commons_transpose, site1="Castril", site2="Pontones")
jcs <- myjaccard(commons_transpose, site1="Castril", site2="Santiago")
jsp <- myjaccard(commons_transpose, site1="Santiago", site2="Pontones")


jshort_long <- myjaccard(tranhs_transpose, site1="Estante", site2="Trashumante")

jces_cts <- myjaccard(ct_transpose, site1="CES", site2="CTS")
jces_pes <- myjaccard(ct_transpose, site1="CES", site2="PES")
jces_pts <- myjaccard(ct_transpose, site1="CES", site2="PTS")
jces_ses <- myjaccard(ct_transpose, site1="CES", site2="SES")
jces_sts <- myjaccard(ct_transpose, site1="CES", site2="STS")

jcts_pes <- myjaccard(ct_transpose, site1="CTS", site2="PES")
jcts_pts <- myjaccard(ct_transpose, site1="CTS", site2="PTS")
jcts_ses <- myjaccard(ct_transpose, site1="CTS", site2="SES")
jcts_sts <- myjaccard(ct_transpose, site1="CTS", site2="STS")

jpes_pts <- myjaccard(ct_transpose, site1="PES", site2="PTS")
jpes_ses <- myjaccard(ct_transpose, site1="PES", site2="SES")
jpes_sts <- myjaccard(ct_transpose, site1="PES", site2="STS")

jpts_ses <- myjaccard(ct_transpose, site1="PTS", site2="SES")
jpts_sts <- myjaccard(ct_transpose, site1="PTS", site2="STS")

jpts_sts <- myjaccard(ct_transpose, site1="SES", site2 = "STS")


summary_commons <- data.frame(
compara = c("Castril - Pontones", "Castril - Santiago", "Santiago - Pontones", 
            "Estantes - Trashumantes",
            "CES - CTS", "CES - PES", "CES - PTS", "CES - SES","CES - STS",
            "CTS - PES", "CTS - PTS", "CTS - SES", "CTS - STS",
            "PES - PTS", "PES - SES", "PES - STS",
            "PTS - SES", "PTS - STS",
            "SES - STS"
            ),
jaccard = c(jcp, jcs, jsp, jshort_long,jces_cts, jces_pes, jces_pts, jces_ses, jces_sts, jcts_pes, jcts_pts, jcts_ses, jcts_sts, jpes_pts, jpes_ses, jpes_sts, jpts_ses, jpts_sts, jpts_sts )
)
```


```{r}
summary_commons |> kbl(digits = c(0,3)) |> 
  kable_styling()
```













